### 2015-09-09 LaTeX/skewnormal.tex
move appendices into supplemental document

@@

\newpage
\section{Introduction}\label{s:intro}
In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
Figure \ref{fig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{plots/ozone-10jul-us.pdf}
  \caption{Ozone values on July 10, 2005}
  \label{fig:ozone-10jul}
\end{figure}

A spatial model for threshold exceedances warrants special consideration and standard spatial methods are likely to perform poorly.
First, because we are interested only in high values, we want to ``let the tail speak for itself''.
That is, if we fit a model to the entire data set, low-to-moderate values would influence the fit of the overall model.
As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
Our inference method will only use data which exceed a threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
However, the Gaussian distribution is light-tailed and symmetric, and therefore may be inappropriate for modeling data which does not share this tail behavior.
Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
Asymptotic dependence/independence (see Section \ref{s:extdep}) are notions which describe how two random variables' dependence behaves as one looks further into the joint tail.
The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the skew-$t$ distribution forms the basis for our model.

@@

Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
This allows us to use Bayesian methods to impute data below the threshold as well as make predictions at unobserved locations.
In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
This partition alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.

@@

\section{Spatial skew processes}\label{s:spatialskew}
Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Azzalini2014}.
A brief review of the additive process by which a skew-$t$ process is created is given here.

@@

\subsection{Censoring to focus on the tails}
As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
Let
\beq\label{Yt}
  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
      Y(\bs) \quad & \delta(\bs) = 1 \\
      T & \delta(\bs) = 0
  \end{array} \right.
\eeq
be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
We impute the censored values as a step in the MCMC algorithm used to fit the model described in Section \ref{s:comp}.

@@

The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
\begin{align}
  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
\end{align}
where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in (\ref{eq:chiskew-t}), $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
In Appendix \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
We estimate $\pi(h)$ through simulation.
% To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
% We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
% For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
% This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.

\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{plots/chi-h.pdf}
  \caption{$\chi$, as a function of distance, for $K = 1, 3, 5$, and $10$ knots.}
  \label{fig:chi}
\end{figure}

@@

We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to the appropriate marginal distribution for the skew-$t$ process.
We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align}
  w^*_{tki} = \Phi^{-1}\left[ \frac{ w_{tki} - \min(\bs_i)}{ \max(\bs_i) - \min(\bs_i) } \right], \quad i = 1, 2
\end{align}
where $\Phi$ is a univariate standard normal density function, and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
Then the transformed knots $\bw^*_{tk} \in \calR^2$.
We use a copula on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align}
where IG is defined as before.
We also use a copula on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
Let
\begin{align}
  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[z_t(\bs)] \right\}
\end{align}
where HN is the distribution function of a half-normal random variable.
The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
\begin{align}
  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align}
where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, \sigma^2_{k})$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, \sigma^2_{tk})$ $\sigma^2_{tk} \sim \text{IG}(a, b)$.
For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.

@@

\section{Simulation study}\label{s:simstudy}
In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model.

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generated data from model (\ref{eq:partition})) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
The sites were generated Uniform$([0, 10] \times [0, 10])$.
We generated data from 4 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=5$ knots
  \item Skew-$t$ marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=5$ knots
  \item Max-stable
  % \item Transformation below $T = q(0.80)$
\end{enumerate}
In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation was $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect was $0.1$.
In the first design, $\sigma^2 = 2$ was used for all days which results in a Gaussian distribution.
For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$ to give a $t$ distribution with 6 degrees of freedom.
For design 1, we set $\lambda = 0$.
For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in (\ref{eq:sitezsig}).
In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ was assumed to be constant across space.
In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
In this design, data have marginal distributions that follow a generalized extreme value distribution with location parameter $1$, scale parameter $1$, and shape parameter $0.2$.
In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
For this setting, we set $\gamma = 0.5$.
% In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
% \begin{align}
%   y = \left\{ \begin{array}{lc}
%     \tilde{y}, \quad & \tilde{y} > T \\[0.5em]
%     T \exp\{\tilde{y} - T\}, \quad & \tilde{y} \le T
%   \end{array}\right.
% \end{align}
% where $T = q(0.80)$ is the 80th sample quantile of the data.
% The final design is included to explore the importance of using a threshold exceedance model when the distribution for the bulk of the data is misspecified.

@@

\appendix
\section{Appendices}
\subsection{MCMC details} \label{a:mcmc}
The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org). Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
When no conjugate prior distribution exists, we use a random walk Metropolis-Hastings update step.
In each Metropolis-Hastings update, we tune the algorithm to give acceptance rates near 0.40.

\subsubsection*{Spatial knot locations}
For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis-Hastings block update.
Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:temporal}) and a random walk bivariate Gaussian candidate distribution
\begin{align*}
	{\bw^*_k}^{(c)} \sim \text{N}({\bw^*_k}^{(r - 1)}, s^2 I_2)
\end{align*}
where ${\bw^*_k}^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
After candidates have been generated for all $K$ knots, the acceptance ratio is
\begin{align*}
  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K} p({\bw^*_k}^{(c)})}{ \prod_{k = 1}^{K} p({\bw^*_k}^{(r - 1)})}\right\}
\end{align*}
where $l$ is the likelihood given in (\ref{eq:hier}), and $p(\cdot)$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be uniform over $\calD$.
The candidate knots are accepted with probability $\min\{R, 1\}$.

\subsubsection*{Spatial random effects}
If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \ref{a:posterior}.
If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis-Hastings update.
Because this model uses $|z_{tk}|$, we generate candidate random effects using the $z^*_{tk}$ (see Section \ref{s:temporal}) and a random walk Gaussian candidate distribution
\begin{align*}
  {z^*_{tk}}^{(c)} \sim \text{N}({z^*_{tk}}^{(r - 1)}, s^2)
\end{align*}
where ${z^*_{tk}}^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
The acceptance ratio is
\begin{align*}
  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}
\end{align*}
where $p[\cdot]$ is the prior taken from the time series given in Section \ref{s:temporal}.
The candidate is accepted with probability $\min\{R, 1\}$.

\subsubsection*{Variance terms}
When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis-Hastings update.
First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
The acceptance ratio is
\begin{align*}
  R = \left\{
    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
    \right\} \times \left\{
    \frac{ l[z_{tk} | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[z_{tk} | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
    \right\} \times \left\{
    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
    \right\} \times \left\{
    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
         { c[ {\sigma^2_{tk}}^{(c)}]}
    \right\}
\end{align*}
where $p[\cdot]$ is the prior either taken from the time series given in Section \ref{s:temporal} or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
The candidate is accepted with probability $\min\{R, 1\}$.

\subsubsection*{Spatial covariance parameters}
We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis-Hastings block update step.
First, we generate a candidate using a random walk Gaussian candidate distribution
\begin{align*}
	\log(\rho)^{(c)} \sim \text{N}(\log(\rho)^{(r - 1)}, s^2)
\end{align*}
where $\log(\rho)^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
Candidates are generated for $\log(\nu)$ and $\gamma$ in a similar fashion.
The acceptance ratio is
\begin{align*}
	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
\end{align*}
All three candidates are accepted with probability $\min\{R, 1\}$.

\subsection{Posterior distributions} \label{a:posterior}
\input{zpost.tex}
\input{betapost.tex}
\input{sigpost.tex}
\input{lambdapost.tex}

\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
Consider a homogeneous spatial Poisson process with intensity $\mu$.
Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
Let $N(A)$ be the number of points in $A$, and let
\begin{align*}
  \mu(A) = \mu |A| = \mu \pi \left(\frac{h}{2}\right)^2 = \lambda h^2.
\end{align*}
Then
\begin{align*}
  P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
                &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
                &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
\end{align*}
which goes to one as $h \rightarrow \infty$.

% Let $N(A)$ be the number of knots in $A$.
% So,
% \begin{align*}
%   P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
% \end{align*}
% Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
% With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
% Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.

% \subsection{Half-normal distribution}
% Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
% Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
% \begin{align}
%   f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
% \end{align}
% When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.

\subsection{Skew-$t$ distribution} \label{a:skewt}
\subsubsection*{Univariate extended skew-$t$ distribution}
We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
\begin{align}
  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left[ (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right]
\end{align}
where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.
In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.

\subsection*{Multivariate skew-$t$ distribution}
If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
\begin{align}
  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
\end{align}
where
\begin{align}
  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^T \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
  \bz &= \bomega^{-1}(\by - \xi)
\end{align}
where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.

>>>>

% cross-referencing appendices
\usepackage{xr}
\externaldocument{supplemental}

@@

\newpage
\section{Introduction}\label{s:intro}
% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
Figure \ref{fig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=0.75\linewidth]{plots/ozone-10jul-us.pdf}
  \caption{Ozone values (ppb) on July 10, 2005}
  \label{fig:ozone-10jul}
\end{figure}

A spatial model for threshold exceedances warrants special consideration and standard spatial methods are likely to perform poorly.
First, because we are interested only in high values, we want to ``let the tail speak for itself''.
That is, if we fit a model to the entire data set, low-to-moderate values would influence the fit of the overall model.
As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
Our inference method will only use data which exceed a threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
However, the Gaussian distribution is light-tailed and symmetric, and therefore may be inappropriate for modeling data which does not share this tail behavior.
Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
Asymptotic dependence/independence (see Section \ref{s:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the skew-$t$ distribution forms the basis for our model.

@@

Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
This allows us to use Bayesian methods to impute data below the threshold as well as make predictions at unobserved locations.
In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
This partitioning alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.

@@

\section{Spatial skew processes}\label{s:spatialskew}
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Azzalini2014}.
A brief review of the additive process by which a skew-$t$ process is created is given here.

@@

\subsection{Censoring to focus on the tails}
As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
% The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
Let
\beq\label{Yt}
  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
      Y(\bs) \quad & \delta(\bs) = 1 \\
      T & \delta(\bs) = 0
  \end{array} \right.
\eeq
be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
We impute the censored values as a step in the MCMC algorithm used to fit the model described in Section \ref{s:comp}.

@@

The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
\begin{align}
  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
\end{align}
where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in (\ref{eq:chiskew-t}), $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
In Appendix \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
In Figure \ref{fig:chi} we estimate $\pi(h)$ through simulation.
% To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
% We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
% For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
% This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.

\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{plots/chi-h.pdf}
  \caption{Extremal dependence measure $\chi(h)$, as a function of distance, $h$, for $K = 1, 3, 5$, and $10$ knots.}
  \label{fig:chi}
\end{figure}

@@

We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to the appropriate marginal distribution for the skew-$t$ process.
We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align}
  w^*_{tki} = \Phi^{-1}\left[ \frac{ w_{tki} - \min(\bs_i)}{ \max(\bs_i) - \min(\bs_i) } \right], \quad i = 1, 2
\end{align}
where $\Phi$ is a univariate standard normal density function and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
Then the transformed knots $\bw^*_{tk} \in \calR^2$.
We use a copula on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align}
where IG is defined as before.
We also use a copula on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
Let
\begin{align}
  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[z_t(\bs)] \right\}
\end{align}
where HN is the distribution function of a half-normal random variable.
The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
\begin{align}
  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align}
where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, \sigma^2_{k})$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, \sigma^2_{tk})$ $\sigma^2_{tk} \sim \text{IG}(a, b)$.
For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.

@@

\section{Simulation study}\label{s:simstudy}
In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model and to compare with Gaussian and max-stable methods.

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generated data from model (\ref{eq:partition}) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
The sites were generated Uniform$([0, 10] \times [0, 10])$.
We generated data from 4 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=1$ knot
  % \item Symmetric-$t$ marginal, $K=5$ knots
  \item Skew-$t$ marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=5$ knots
  \item Max-stable
  % \item Transformation below $T = q(0.80)$
\end{enumerate}
In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation was $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect was $0.1$.
In the first design, $\sigma^2 = 2$ was used for all days which results in a Gaussian distribution.
For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$ to give a $t$ distribution with 6 degrees of freedom.
For design 1, we set $\lambda = 0$.
For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in (\ref{eq:sitezsig}).
In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ was assumed to be constant across space.
In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
In this design, data have marginal distributions that follow a generalized extreme value distribution with location parameter $1$, scale parameter $1$, and shape parameter $0.2$.
In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
For this setting, $\gamma_{\text{RS}} \in (0, 1)$, is similar to $\gamma$ in that it controls the relative strength of the nugget effect.
We set $\gamma_{\text{RS}} = 0.5$, which represents moderate spatial dependence.
% In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
% \begin{align}
%   y = \left\{ \begin{array}{lc}
%     \tilde{y}, \quad & \tilde{y} > T \\[0.5em]
%     T \exp\{\tilde{y} - T\}, \quad & \tilde{y} \le T
%   \end{array}\right.
% \end{align}
% where $T = q(0.80)$ is the 80th sample quantile of the data.
% The final design is included to explore the importance of using a threshold exceedance model when the distribution for the bulk of the data is misspecified.

@@
